---
title: "Intro to Eartquake Tools"
author: "Tom Cooper"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to Eartquake Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The eqtools package makes it easy for  users to explore the  U.S. National Oceanographic and Atmospheric Administration (NOAA) dataset on significant earthquakes around the world. This dataset contains information about 5,933 earthquakes over an approximately 4,000 year time span.

### Downloading dataset

The datset is available from the [NOAA Website](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1)
The entire dataset can be download by clicking on the link: "Download entire significant earthquake data file in tab-delimited format."

### What Eqtools includes

- Functions to clean the data to prepare for 
- Geoms to create timeline plots
- Functions to create interactive html maps 

### Data: NOAA significant earthquake data 

```{r, echo = TRUE, eval = TRUE, load_data}
     eqdata <- readr::read_delim(  system.file("extdata",
                          "signif-july-2017.txt.gz",
			   package = "eqtools"),delim = "\t")
    dim(eqdata)
```

### Cleaning tools

```{r, echo = TRUE, eval = TRUE, clean_data}
    library(eqtools)
    str(eqdata$LATITUDE)
    eq_clean <- eq_clean_data(eqdata)
    str(eq_clean$LATITUDE)
```

```{r, echo = TRUE, eval = TRUE, clean_location}
    eqdata$LOCATION_NAME[1]
    eq_clean <- eq_clean_location(eqdata)
    eq_clean$LOCATION_NAME[1]
```

### Plotting Tools
library(dplyr)
```{r, echo = TRUE, eval = TRUE, timeline_plot}
    eq_us_mex <- dplyr::mutate(eq_clean)
    aplot <- ggplot(eq_us_mex,aes(x = YEAR, y = 1,
         xmin = minYear,xmax = maxYear,
         )) +
  geom_timeline(aes(size = EQ_PRIMARY))

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
